Multi-level modeling with Openflexo/\FML

### 2021-09-24 10-conclusions.tex
Finishing related work. A word on constraint is missing.

We fulfill all the challenge requirements and we propose a tooling that
demonstrates the usability of our solution. We developed an ad hoc solution,
including models and metamodels, thanks to the flexibility provided by the
metametamodeling infrastructure offered by Openflexo. \noteSM{This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object. In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements. This, together with the FML core support for (level agnostic) classical linguistic instantiation permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.}

>>>>

We fulfill all the challenge requirements and we propose a tooling that
demonstrates the usability of our solution. We developed an ad hoc solution,
including models and metamodels, thanks to the flexibility provided by the
metametamodeling infrastructure offered by Openflexo. \textcolor{red}{This infrastructure helped us to overcome the limitations of the strict modeling paradigm mentioned in Section \ref{sec:introduction}, this is, the lack of support for different forms of classifications and for the duality type/object. In that sense, our solution seamlessly integrates the type-object pattern which allows us to dynamically create different \emph{instances} of model elements (e.g., \emph{TaskType}, among others) and use them for typing other model elements. This, together with the FML core support for (level agnostic) classical linguistic instantiation permitted us to merge linguistic and ontological instantiation in order to provide a satisfactory solution to the Multi-Level Process challenge.}

### 2021-09-24 9-relatedworks.tex
Finishing related work. A word on constraint is missing.

More similar to us, in the submission \parencite{deeptelos2019}, the author uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge. As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency. However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} \textcolor{red}{which the authors use to simulate potency. Models in DeepTelos are organized in (tree-like) module hierarchies, where submodules can see all the concepts defined in parent modules. The authors use this module system in order to \emph{organize} their solution for the Multi-level process challenge. Concretely they created a hierarchy of modules such as the top module contains base concepts and formulas required for multi-level modeling (e.g., the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19. This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules}.

@@

Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge. DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented). Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain. Wrappers are used in order to reuse tasks in processes. Their solution does not use inheritance, as it is not supported by the framework. It does not use explicitly separated models either.

>>>>

More similar to us, in the submission \parencite{deeptelos2019}, the author uses DeepTelos \parencite{deeptelos2016}, an extension of the Telos language \parencite{telos1990} in order to solve the process challenge. As with Openflexo/\FML, they do not use explicit (numbered) levels nor potency. However, unlike Openflexo, DeepTelos integrates a multi-level modeling specific construct similar to the PowerType \parencite{atkinson2001essence} pattern they call \emph{most general instances} \textcolor{red}{which the authors use to simulate potency. Models in DeepTelos are organized in (tree-like) module hierarchies, where submodules can see all the concepts defined in parent modules. The authors use this module system in order to \emph{organize} their solution for the Multi-level process challenge. Concretely they created a hierarchy of modules such as the top module contains base concepts and formulas required for multi-level modeling (e.g., the support for \emph{most general instances}), and a sub-module contains process definitions fulfilling requirements P1 to P19. This sub-module contains in turn a sub-module representing the coding process type whereas concrete coding processes are represented in subsequent sub-modules}. There are two main differences between the DeepTelos and Openflexo/\FML: 1) we use the type-object pattern instead of the powertype pattern; 2) DeepTelos reify their support for multi-level modeling while ours remains ad-hoc.

@@

Similarly, \cite{dmla2019} use the Dynamic Multi-Layer Algebra (DMLA) \parencite{dmla2017} in order to solve the process challenge. DMLA is a level-blind modeling framework which is fully customizable (\eg different types of instantiation may be implemented) \textcolor{red}{and includes support for deep characterization (instantiation may refer to any other element disregarding hierarchy levels and a sort of potency in the form of fluid metamodeling at the entity level exists).}

Their proposed solution separates the process challenge in two separate domains, namely, the task definition domain and the process definition domain. Wrappers are used in order to reuse tasks in processes. Their solution does not use inheritance, as it is not supported by the framework. It does not use explicitly separated models either.
